
s---
title: "Data Cleaning"
author: "Steven M. Karceski"
date: "November 5, 2018"
output: html_document
---

```{r setup, message = F, warning = F}

rm(list = ls())

library(tidyverse)
library(sjPlot)
library(pander)

data <- read_csv("FINAL_SOC403_AU2018_Survey.csv")

```

The first chunk of code (above) loads two package libraries that are necessary to run the code. If you haven't already done so, install the packages using the function `install.packages("package_name")`. The second part of the chunck is reading in our survey data with the `read_csv()` function (part of the `tidyverse::` package we loaded before. 

The data need to be downloaded from the survey page. You can do this by clicking on the three vertical dots on the Responses page and select "Download responses (.csv)". Save the file to the same folder you saved this file and the code above should work. 

As I read in the data ("  `PreliminaryResponseData.csv`") I create an object called `data` with remains in our environment (you should see the object name, the number of observations, and the number of variables). 

## Renaming Columns

The columns in the data are named with the literal text of each question. We need to change the names to shorter, but still intutive labels. We also need to create a codebook that matches each survey question to the new names. It should also define each variable we create later on. This code uses the `rename()` function from the `dplyr` package, part of the `tidyverse` library loaded earlier. When creating new column names, avoid using spaces in the names; R handles column names better this way. 

```{r rename}

data <- data %>%  # create another data object so that we can easily go back to the original
    rename("TimeStamp" = "Timestamp",
           "UW_Undergrad" = "Are you currently enrolled as an undergraduate student in Washington State?",
           "Employed" = "Are you currently employed (including paid internships)?",
           "Employed_Multiple" = "Are you currently working more than one job?",
           "Employed_Part" = "Are you employed full or part time?",
           "Employed_Self" = "Are you self-employed?",
           "Job_Sector" = "What job sector do you primarily work in?",
           "Employed_Campus" = "Are you currently working ON campus or OFF campus?",
           "Employed_Seattle" = "Do you work within the City of Seattle?",
           "IW_Notice" = "How many days prior notice do you typically receive regarding your upcoming work schedule?",
           "IW_On_Call" = "How frequently do you work on-call shifts? (Employees who work on-call are expected to be available during the shift at any time of day or night, usually with short notice, to carry out their working duties)",
           "IW_Hrs_Compare" = "How do the weekly hours you work compare to hours originally indicated upon hiring?",
           "IW_Clopening" = "Have you ever worked a \"clopening\" shift? (Clopening is when the employee has to work until closing time on a given day, and then they open on the next day.)",
           "IW_Control_When" = "How much influence do you have over when you are scheduled to work?",
           "IW_Control_Amount" = "How much influence do you have over the amount of hours you work?",
           "IW_Late_Pay" = "Have you ever received a paycheck later than expected date in the last couple of months?",
           "IW_Lower_Pay" = "Have you ever been paid less than the agreed-upon wage in the last couple of months?",
           "IW_Overtime" = "Do you get compensated financially for working more than expected (such as bonuses or time and a half for holidays/overtime)?",
           "IW_Tips" = "Do you receive tips or commission for your job?",
           "IW_Tips_PCT" = "Approximately what percentage of your earnings are from commissions or tips per month?",
           "IW_Seasonal" = "Is your job seasonal?",
           "IW_Contract" = "Do you work under a contract that employs you under a specific time frame?",
           "IW_Benefits" = "Check all benefits provided by your job. (Don't check any if none apply.)",
           "IW_Tenure" = "How long do you expect to be working for your current employer?",
           "IW_Advancement" = "Do you think it's likely you might be promoted/ advance in your current job?",
           "IW_Decline_Shift" = "When asked to pick up an extra shift, I feel comfortable declining.",
           "IW_Understanding" = "My manager is understanding of my commitments to school when it comes to setting the schedule.",
           "IW_Conditions" = "In general, I am satisfied with my current work conditions.",
           "IW_SSSO" = "Have you heard of the Seattle Secure Scheduling Ordinance?",
           "F_Income" = "What is your average monthly income?",
           "F_Resp_Pay" = "Are you individually responsible to pay for the following? (Check all that apply)",
           "F_Rent" = "Rent",
           "F_Groceries" = "Groceries",
           "F_Transit" = "Transportation (Gas, car insurance, bus, etc)",
           "F_Supplies" = "School supplies (textbook, equipment, etc)",
           "F_400" = "If you had to come up with $400 to cover an emergency expense, how would you pay for it? (choose your best option)",
           "F_Loans" = "In the last year, have you used any of the following types of loans? (check all that apply)",
           "F_Support" = "How much financial support do you generally receive from family, friends, guardians, or relatives?",
           "F_Sup_Source" = "Do you receive support from: (check all that apply)",
           "A_School" = "What school are you currently attending?",
           "A_GPA" = "What is your current cumulative GPA?",
           "A_Study" = "How long do you spend studying per week (excluding in-class time)?",
           "A_Class_Attend" = "How often do you attend your classes?",
           "A_Class_Late" = "How often are you late to class?",
           "A_Class_Attn" = "In general, are you able to pay attention during the classes you attend?",
           "A_Time_Off" = "Have you ever taken time off of school to care for someone, for financial reasons, or to work?",
           "A_Extra" = "Do you participate in any of the following extracurricular activities? (Select all that apply)",
           "A_Extra_Hrs" = "If applicable, how many hours a week do you spend participating in your extracurricular activities?",
           "Health" = "How would you rate your general health?",
           "Overwhelmed" = "I often feel overwhelmed.",
           "Sat_Life" = "I am generally satisfied with my life.",
           "Sat_Friends" = "I am satisfied with the amount of time I spend with my friends and family (hanging out, visiting, etc...).",
           "Sat_Myself" = "I am satisfied with the amount of time I spend on myself (relaxing, hobbies, personal time, etc...).",
           "Sat_Sleep" = "I am satisfied with the amount of sleep I get.",
           "Sleep" = "In general, how many hours of sleep do you get per day?",
           "A_Grade" = "What is your class standing in college?",
           "Residence" = "Where do you currently reside?",
           "Reside_With" = "Who do you currently reside with?",
           "In_State" = "Are you considered an in-state or out-of-state student for tuition purpose?",
           "Major" = "What is your major? (if you declared; if you have not, what is your intended major. Example: Sociology)",
           "Race_Ethnicity" = "How would you describe your race/ethnicity? (Please select one that you most identify with)",
           "Citizen" = "Are you a citizen of the United States?",
           "Native_Born" = "Were you born in the United States?",
           "Reside_Seattle" = "Do you currently live within the Seattle city limits?",
           "Gender" = "What is your gender identity?",
           "Gender_Trans" = "Do you identify as transgender?",
           "Sexuality" = "Do you consider yourself to be:",
           "Marital_Stat" = "What is your marital status?",
           "Income_Source" = "What is your main source of income (the source that you receive the most support from)?",
           "Dependents" = "Are there one or more individuals reliant on you for caregiving and/or financial support?",
           "Parental_Income" = "What is your parent/guardian household income?",
           "Parental_Edu" = "What is the highest level of education that either of your parents or guardians received?",
           "Parental_Citizen" = "Are your parent(s)/guardian(s) U.S. citizens?",
           "Follow_Up" = "Are you interested in participating in a follow up interview regarding insecure work?",
           "Folow_Up _Email" = "If you answered \"yes\" in the last question, make sure to also include your email below. We will reach out to you to schedule an interview shortly!",
           "Gift_Card" = "Please leave your email if you interested in entering the drawing of three $25 amazon gift card here.")

write_csv(data, "SurveyData_Renamed.csv")

```

## Creating Variables

Even after renaming variables, the survey data are not currently in a form that is very useful for us. In order for it to be useful, we need to create variables so we can look at relationships between other variables. For example, the question on how much advance notice respondents get for their schedule can be made into a dichotomous variable (takes one of two values). We can do this by giving the value of 1 to responses that indicate they receive less than two weeks notice and a 0 to all other responses. I will demonstrate this below using the`mutate()` function in the `dplyr` package. 

```{r coding}

# the code below creates a new variable (Notice_Less2wk) using the ifelse() function. This is saying, if the response to the 
# Notice variable is 7-13 days OR 0-6 days, then the new variable takes a value of 1. Otherwise it takes a value of zero. 
# ifelse(logical statement, value if True, value if false) 
# for help with using any function in R, type ?function_name() in the console and this will bring up the appropriate help file.  


NewData <- data %>%  
    mutate(
        # section 2
        Student = ifelse(UW_Undergrad != "No", 1, 0),
        # section 3
        Employed = ifelse(Employed == "Yes", 1, 0),
        Employed_one = ifelse(Employed_Multiple == "No", 1, 0),
        # section 4
        Employed_full = ifelse(Employed_Part == "Full-time (work on average 35 or more hours per week)", 1, 0),
        Employed_self = ifelse(Employed_Self == "Yes", 1, 0),
        # Job_Sector
        Employed_oncampus = ifelse( Employed_Campus == "ON campus", 1, 0),
        Seattlework = ifelse(Employed_Seattle == "Yes", 1, 0),
        # section 5
        Notice_Less2wk = ifelse(IW_Notice %in% c("7 - 13 days", "0 - 6 days"), 1, 0),
        Notice_Less1wk = ifelse(IW_Notice == "0 - 6 days", 1, 0),
        No_oncall = ifelse(IW_On_Call == "Never" , 1, 0),
        Hours_same = ifelse(IW_Hrs_Compare == "Roughly the same", 1, 0),
        Clopen = ifelse(IW_Clopening== "Yes", 1, 0),
        Control_when = ifelse(IW_Control_When == "Complete", 1, 0),
        Control_hours = ifelse(IW_Control_Amount == "Complete", 1, 0),
        Latepay = ifelse(IW_Late_Pay == "Yes", 1, 0),
        Lesspay = ifelse(IW_Lower_Pay == "Yes", 1, 0),
        Overtime_compensate = ifelse(IW_Overtime == "Yes", 1, 0),
        Job_Notip = ifelse(IW_Tips == "No", 1, 0),
        # Only use the response under yes/no question here for tip because we are only interested in whether they 
        # receive tip or not, and those who do represent a really small percentage. 
        Job_seasonal = ifelse(IW_Seasonal == "Yes", 1, 0),
        Job_contract = ifelse(IW_Contract == "Yes", 1, 0),
        # Benefit: check all that apply
        Insurance_ben = ifelse(str_detect(IW_Benefits,"Insurance") == "TRUE", 1, 0),
        Vacation = ifelse(str_detect(IW_Benefits,"Vacation")== "TRUE", 1, 0),
        PTO = ifelse(str_detect(IW_Benefits,"PTO") == "TRUE", 1, 0),
        Work_more_1yr = ifelse(IW_Tenure %in% c("1 - 2 years", "More than 2 years"), 1, 0),
        Job_promote = ifelse(IW_Advancement == "Yes", 1, 0),
        Decline_comfy = ifelse(IW_Decline_Shift %in% c("4", "5"), 1, 0),
        Manager_good = ifelse(IW_Understanding %in% c("4", "5"), 1, 0),
        Work_satisfy = ifelse(IW_Conditions %in% c("4", "5"), 1, 0),
        KnowOrdinance = ifelse(IW_SSSO == "Yes", 1, 0),
        MonthEarnLess999 = ifelse(F_Income == "less than $999", 1, 0),
        # section 6
        # Individual responsibility: check all that apply
        payRent = ifelse(str_detect(F_Resp_Pay,"Rent") == "TRUE", 1, 0),
        payGrocery = ifelse(str_detect(F_Resp_Pay,"Groceries") == "TRUE", 1, 0),
        payClothes = ifelse(str_detect(F_Resp_Pay,"Clothes") == "TRUE", 1, 0),
        payInsurance = ifelse(str_detect(F_Resp_Pay,"Insurance") == "TRUE", 1, 0),
        payUtility = ifelse(str_detect(F_Resp_Pay,"Utility ") == "TRUE", 1, 0),
        payPhonebill = ifelse(str_detect(F_Resp_Pay,"Phone") == "TRUE", 1, 0),
        payTransport = ifelse(str_detect(F_Resp_Pay,"Transportation") == "TRUE", 1, 0),
        paySchool = ifelse(str_detect(F_Resp_Pay,"School") == "TRUE", 1, 0),
        Rent_Nodifficulty = ifelse(F_Rent %in% c("Never", "Rarely"), 1, 0),
        Grocery_Nodifficulty = ifelse(F_Groceries %in% c("Never", "Rarely"), 1, 0),
        Transport_Nodifficulty = ifelse(F_Transit %in% c("Never", "Rarely"), 1, 0),
        School_Nodifficulty = ifelse(F_Supplies %in% c("Never", "Rarely"), 1, 0),
        Rent_Difficulty = ifelse(F_Rent %in% c("Very Often", "Always"), 1, 0),
        Grocery_Difficulty = ifelse(F_Groceries %in% c("Very Often", "Always"), 1, 0),
        Transport_Difficulty = ifelse(F_Transit %in% c("Very Often", "Always"), 1, 0),
        School_Difficulty = ifelse(F_Supplies %in% c("Very Often", "Always"), 1, 0),
        Have400 = ifelse(F_400 == "Have it in saving account", 1, 0),
        # Types of loans: check all that apply
        paydayLoan = ifelse(str_detect(F_Loans, "Payday") == "TRUE", 1, 0),
        CreditcardLoan = ifelse(str_detect(F_Loans, "Credit") == "TRUE", 1, 0),
        ffLoan= ifelse(str_detect(F_Loans, "friends or family") == "TRUE", 1, 0),
        bankLoan = ifelse(str_detect(F_Loans, "Bank") == "TRUE", 1, 0),
        mostOrAllsupport = ifelse(F_Support %in% c("All my expenses are covered", "Most"), 1, 0),
        # Support: check all that apply
        Scholarship = ifelse(str_detect(F_Sup_Source, "Scholarship") == "TRUE", 1, 0),
        Grant = ifelse(str_detect(F_Sup_Source, "Grant") == "TRUE", 1, 0),
        studentLoan = ifelse(str_detect(F_Sup_Source, "Student") == "TRUE", 1, 0),
        Stipend = ifelse(str_detect(F_Sup_Source, "Stipend") == "TRUE", 1, 0),
        Fellowship = ifelse(str_detect(F_Sup_Source, "Fellowship") == "TRUE", 1, 0),
        # section 7
        School_uw = ifelse(A_School == "University of Washington - Seattle", 1, 0),
        Insuff_studytime = ifelse(A_Study %in% c("0 to 5", "6 to 10"), 1, 0),
        Time_off = ifelse(A_Time_Off == "Yes", 1, 0),
        Less_Extracurriculars = ifelse(A_Extra_Hrs %in% c("0 - 2hours", "3 - 5hours"), 1, 0), 
        Class_freq = ifelse(A_Class_Attend %in% c("4","5"), 1, 0),  
        Late_freq = ifelse(A_Class_Late %in% c("4", "5"), 1, 0),
        Attn_freq = ifelse(A_Class_Attn %in% c("4", "5"), 1, 0),
        # Extracurricular activity: check all that apply
        RSO = ifelse(str_detect(A_Extra, "RSO")== "TRUE", 1, 0),
        Sport = ifelse(str_detect(A_Extra, "Sports")== "TRUE", 1, 0),
        Music = ifelse(str_detect(A_Extra, "Music")== "TRUE", 1, 0),
        Religion = ifelse(str_detect(A_Extra, "Religious")== "TRUE", 1, 0),
        Volunteer = ifelse(str_detect(A_Extra, "Volunteer")== "TRUE", 1, 0),
        Internship = ifelse(str_detect(A_Extra, "Internship")== "TRUE", 1, 0),
        # section 8
        Healthy = ifelse(Health %in% c("Good", "Excellent"), 1, 0),
        Overwhelmed = ifelse(Overwhelmed %in% c("4", "5"), 1, 0),
        Satified_life = ifelse(Sat_Life %in% c("4", "5"), 1, 0),
        Satisfied_friends = ifelse(Sat_Friends %in% c("4", "5"), 1, 0),
        Satisfied_myself = ifelse(Sat_Myself %in% c("4", "5"), 1, 0),
        Satisfied_sleep = ifelse(Sat_Sleep %in% c("4", "5"), 1, 0),
        Suff_sleep = ifelse(Sleep %in% c("7-8 hours", "9 or more hours"), 1, 0),
        # No check all that apply
        # section 9
        Have_citizenship = ifelse(Citizen == "Yes", 1, 0),
        US_born = ifelse(Native_Born == "Yes", 1, 0),
        Reside_seattle = ifelse(Reside_Seattle == "Yes", 1, 0),
        Rside_noparents = ifelse(Reside_With %in% c("By myself", "With roommates (may include partner", "With partner (no other
                                                    roommates)"), 1, 0),
        Instate_tuition = ifelse(In_State == "In-state", 1, 0),
        # this incorrectly mixes gender with sexulity - invalid
        # Straight_gender = ifelse(Gender %in% c("Female", "Male"), 1, 0), 
        Transgender = ifelse(Gender_Trans == "Yes", 1, 0),
        Straight = ifelse(Sexuality == "Straight", 1, 0),
        Single = ifelse(Marital_Stat == "Single, never married", 1, 0),
        Sup_family = ifelse(Income_Source == "Family and/or Friends", 1, 0),
        No_dependents = ifelse(Dependents == "No", 1, 0),
        Low_income = ifelse(Parental_Income %in% c("Less than $35,000", "$35,000 to $69,000"), 1, 0),
        Not_Educated_parents = ifelse(Parental_Edu %in% c("Less than high school", "High school", "Associates/2-year degree"),
                                      1, 0),
        Parent_Bach = ifelse(Parental_Edu %in% c("4-year degree", 
                                                  "Graduate (Master / Ph.D.) or Professional Degree (M.D. / J.D.)"), 1, 0),
        Citizen_parents = ifelse(Parental_Citizen == "Yes", 1, 0),
        A_GPA = ifelse(A_GPA > 4, NA, A_GPA))

# remove duplicates and non-students: 
NewData <- NewData %>% distinct() %>% filter(Student == 1) 

# this will write a new csv file 
write_csv(NewData, "Full_Survey_Data.csv") 

```

### Exploting the associations between our variables

In the next section, I focus on one of our insecure work variables, `Work_satisfy`, and look at associations between it and the demographic/SES variables, as well as several of the outcome variables. The code I have below can be used by each of you to check explore you own assigned variables. Play around with it, don't be afraid to make mistakes, and feel free to email me/post in the slack if you have any questions. 

To start, the first bit of code shows how you can easily get a count for each response for a certain variable. Here I do it for race, but it can easily be replicated for other variables by inputting the name of some other variable in place of `Race_Ethnicity`. 

```{r determinants, include = T}

# let's find out how many observations we have ggot each race/ethnicity category
NewData %>% 
    group_by(Race_Ethnicity) %>% 
    summarize(Count = n()) %>% 
    pander()

NewData %>% 
    filter(Employed == 1) %>% # here we limit respondents to just those who are employed
    # the variable for which we want to see the association with work satisfactionis is Race_ethnicity
    group_by(Race_Ethnicity) %>% 
    summarize(Work_satisfy = mean(Work_satisfy), # the mean of the variable for each value Work_satisfy can take
              Count = n()) %>%  
    pander() # print a table

# I can do the same thing for gender:
NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Gender) %>% 
    summarize(Work_satisfy = mean(Work_satisfy),
              Count = n()) %>%   
    pander() 

# And for sexuality
NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Sexuality) %>% 
    summarize(Work_satisfy = mean(Work_satisfy),
              Count = n()) %>%   
    pander() 

# Let's look at some of the SES categories: 
NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Parental_Edu) %>% # group by parental educatioon (you can also use the new variable) 
    summarize(Work_satisfy = mean(Work_satisfy),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Parental_Income) %>% # group by parental income
    summarize(Work_satisfy = mean(Work_satisfy),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Citizen) %>% # group by citizenship
    summarize(Work_satisfy = mean(Work_satisfy),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Income_Source) %>% # group by income source
    summarize(Work_satisfy = mean(Work_satisfy),
              Count = n()) %>%   
    pander() 

```
 
 ### Outcomes
 
Now let's take a look at how the insecure work varialbe is associated with various variables. What you can do to run this for your own variable is search (ctrl + F) and replace `Work_satisfy` with your variable(s) name. 
 

```{r outcomes}


NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Work_satisfy) %>% 
    summarize(Healthy = mean(Healthy),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Work_satisfy) %>% 
    summarize(Overwhelmed = mean(Overwhelmed),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Work_satisfy) %>% 
    summarize(Satified_life = mean(Satified_life),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Work_satisfy) %>% 
    summarize(Satisfied_sleep = mean(Satisfied_sleep),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Work_satisfy) %>% 
    summarize(A_GPA = mean(A_GPA, na.rm = T),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Work_satisfy) %>% 
    summarize(Rent_Difficulty = mean(Rent_Difficulty),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Work_satisfy) %>% 
    summarize(Grocery_Difficulty = mean(Grocery_Difficulty),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Work_satisfy) %>% 
    summarize(Transport_Difficulty = mean(Transport_Difficulty),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Work_satisfy) %>% 
    summarize(Have400 = mean(Have400),
              Count = n()) %>%   
    pander() 


```

### What is causing Insecure work? 
## INSECURE WORK: LATE PAY 
NewData %>% 
    filter(Employed == 1) %>% 
    filter(Latepay == 1) %>%
    summarize(Race_Ethnicity) %>%
              Count = n()) %>%  
    pander()
## Excluding data which has a too small sample size, race does not play an important factor in deciding which group gets late pays. 
NewData %>% 
    filter(Employed == 1) %>%
    group_by(Race_Ethnicity) %>%
    summarize(Lesspay = mean(Lesspay),
            Count = n()) %>%
    pander()
    
NewData %>% 
    filter(Employed == 1) %>%
    group_by(Sexuality) %>%
    summarize(Latepay = mean(Latepay), Lesspay = mean(Lesspay),
            Count = n()) %>%
    pander()
    
NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Parental_Edu) %>% 
    summarize(Latepay = mean(Latepay), Lesspay = mean(Lesspay),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Parental_Income) %>% # group by parental income
    summarize(Latepay = mean(Latepay),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Citizen) %>% # group by citizenship
    summarize(Latepay = mean(Latepay),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Income_Source) %>% # group by income source
    summarize(Latepay = mean(Latepay),
              Count = n()) %>%   
    pander() 




## OUTCOME of INSECURE WORK - LATE PAY

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(Healthy = mean(Healthy),Overwhelmed = mean(Overwhelmed),Satified_life = mean(Satified_life),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Latepay) %>% 
    summarize(Overwhelmed = mean(Overwhelmed),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Latepay) %>% 
    summarize(Satified_life = mean(Satified_life),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Latepay) %>% 
    summarize(Satisfied_sleep = mean(Satisfied_sleep),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Latepay,Lesspay) %>% 
    summarize(A_GPA = mean(A_GPA, na.rm = T),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Latepay, Lesspay) %>% 
    summarize(Rent_Difficulty = mean(Rent_Difficulty),Grocery_Difficulty = mean(Grocery_Difficulty),
              Count = n()) %>%   
    pander() 
    
NewData %>% 
    filter(Employed == 1) %>%
    filter(Latepay == 1) %>%
    group_by(Gender) %>%
    summarize(Rent_Difficulty = mean(Rent_Difficulty),Grocery_Difficulty = mean(Grocery_Difficulty),
            Count = n()) %>%
    pander()
    

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Latepay) %>% 
    summarize(Grocery_Difficulty = mean(Grocery_Difficulty),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Latepay) %>% 
    summarize(Transport_Difficulty = mean(Transport_Difficulty),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Latepay) %>% 
    summarize(Have400 = mean(Have400),
              Count = n()) %>%   
    pander() 
## LATE PAY 
## There was no particularly outstanding variable that is causing insecure work, late pay. It seems like there is no outstanding variable that is contributing to individual's likely to be in insecure work. 
## LATE PAY OUTCOME 
## However, the outcomes of Insercure work, late pay, has a significant differences than that of individuals who have not experienced late pay. 
##Individuals who have experienced late pay generally have also experienced difficulty paying rent, grocery, bills, and transportation 


### What is causing LESS PAY? 
### INSECURE WORK : LESS PAY 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Race_Ethnicity) %>%
    summarize(Lesspay = mean(Lesspay), 
              Count = n()) %>%  
    pander()
    
NewData %>% 
    filter(Employed == 1) %>%
    group_by(Gender) %>%
    summarize(Lesspay = mean(Lesspay),
            Count = n()) %>%
    pander()
    
NewData %>% 
    filter(Employed == 1) %>%
    group_by(Sexuality) %>%
    summarize(Lesspay = mean(Lesspay),
            Count = n()) %>%
    pander()
    
NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Parental_Edu) %>% 
    summarize(Lesspay= mean(Lesspay),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Parental_Income) %>% # group by parental income
    summarize(Lesspay = mean(Lesspay),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Citizen) %>% # group by citizenship
    summarize(Lesspay = mean(Lesspay),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Income_Source) %>% # group by income source
    summarize(Lesspay= mean(Lesspay),
              Count = n()) %>%   
    pander() 


## what is the outcome of less pay? 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(Healthy = mean(Healthy),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(Overwhelmed = mean(Overwhelmed),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(Satified_life = mean(Satified_life),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(Satisfied_sleep = mean(Satisfied_sleep),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(A_GPA = mean(A_GPA, na.rm = T),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(Rent_Difficulty = mean(Rent_Difficulty),
              Count = n()) %>%   
    pander() 
## SIGNIFICANT : worth bringing attention to 
NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(Grocery_Difficulty = mean(Grocery_Difficulty),
              Count = n()) %>%   
    pander() 
## significant higher rate of difficulty reported by those who have experienced lower pay than those who have not. 
NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(Transport_Difficulty = mean(Transport_Difficulty),
              Count = n()) %>%   
    pander() 

NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(Have400 = mean(Have400),
              Count = n()) %>%   
    pander()
  
  
summary(glm(Lesspay ~ Grocery_Difficulty + Transport_Difficulty,
    data = NewData, family = "binomial"))
    
    
NewData %>% 
    filter(Employed == 1) %>%
    group_by(Race_Ethnicity) %>%
    summarize(Lesspay = mean(Lesspay), Latepay = mean(Latepay),
            Count = n()) %>%
    pander()
    
    
NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(Rent_Difficulty = mean(Rent_Difficulty),Grocery_Difficulty = mean(Grocery_Difficulty),
              Count = n()) %>%   
    pander() 
    
NewData %>% 
    filter(Employed == 1) %>% 
    group_by(Lesspay) %>% 
    summarize(Healthy = mean(Healthy),Overwhelmed = mean(Overwhelmed),Satified_life = mean(Satified_life),
              Count = n()) %>%   
    pander() 

```{r}
NewData %>% 
    filter(Employed == 1) %>%
    group_by(Decline_comfy) %>%
    summarize(Rent_Difficulty = mean(Rent_Difficulty),
              Grocery_Difficulty = mean(Grocery_Difficulty),
              Have400 = mean(Have400),
              Insuff_studytime = mean(Insuff_studytime),
              Healthy = mean(Healthy),
              Overwhelmed = mean(Overwhelmed),
              Satisfied_life = mean(Satified_life),
              Satisfied_sleep = mean(Satisfied_life),
              A_GPA = mean(A_GPA,na.rm = T),
            Count = n()) %>%
    pander()

S

```